---
title: "Group 2 reverse engineering"
author: "Ijeoma Opara"
date: "October 17, 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---//
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, **Nearly 250 women have been**

**fatally shot by police since 2015.**

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

fatal_police_shootings<- read.csv("data/fatal-police-shootings-data.csv") |> clean_names()
agencies<-read.csv("data/fatal-police-shootings-agencies.csv") |> clean_names()

```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: Of the 247 women fatally shot, 48 were Black and seven of those were unarmed.
-   **Analysis summary**: Although the data here said 48 Black women were killed, out of which 7 were unarmed, the current analysis show that the figure has increased from 48 because the data was updated.

```{r}
# Put code to reverse engineer sentence here
fatal_police_shootings |> 
  group_by(gender, race) |> 
  summarize(count=n())

fatal_police_shootings |> 
  group_by(gender)
  summarize
# Display results of code below this codeblock

```

### Sentence 2

-   **Sentence text**: About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed.
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

fatal_police_shootings |> 
  group_by(signs_of_mental_illness, gender) |>
  summarize(total=n()
  ) |>
   arrange(desc(total))

# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: Since 2015, police have killed 26 unarmed women, including Taylor. Of those, 14 were White, seven were Black, four were Hispanic and one’s race was unknown.
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
unarmed_women_killed<-fatal_police_shootings |>
  filter(armed=="unarmed") |>
  group_by(armed, gender, race) |>
  summarize(count=n())
  

unarmed_women_killed



# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: Of the 89 women killed at residences where they lived or often stayed, 12 encounters began as did Taylor’s — with a warrant to conduct a search or make an arrest in an investigation.
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here


# Display results of code below this codeblock

```

### Sentence 5

-   The average age of women and men killed by police was 37, and about one-third were 25 to 34.
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
average_age_of_women <-fatal_police_shootings |>
  group_by(gender) |>
  summarize(
    average_age_of_women =mean(age, na.rm=TRUE)
  )

# Display results of code below this codeblock

```

-30-
